---
title: "ESA watershed figs"
author: "PR & AMP"
date: "2023-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      message = F, 
                      warning = F)

 # load packages
 require(pacman)
 pacman::p_load(tidyverse, # keep things tidy!!
                googlesheets4, # read_sheet
                googledrive, # google drive access
                janitor,
                ggplot2, #plotting
                PNWColors ) # pretty colors 
```

### Load Data

```{r Import L1 data}

l1_modeloutput_all <- read_csv("../230718_output_datset_L1.csv")

```

### Calculate metrics 

```{r}

l1_model_statsums <- l1_modeloutput_all %>%
  select(-dates) %>%
  group_by(short_scenario, scenario, scenario_group, description, scenario_raw, climate, land_use, hydrology, slope, percent, severity) %>%
  dplyr::summarise(across(everything(), list(mean = mean, median = median, sd = sd), na.rm=TRUE))

```

l1_modeloutput_all %>%
 filter(scenario_group %in% c("Climate", "Base Case", "Land Use")) %>%
    ggplot() +
    geom_line(aes(x = dates, y=doc_mg_l_fire, color= description, linetype = description), position= position_jitterdodge()) +   
    ylab("Change in DOC (%)")+
    theme_classic() +
    theme(legend.position = "bottom")+
    guides(color= guide_legend(nrow= 2))
    
    
### ESA plots 
Plots to make are:
DOC
NO3
TSS

climate, vegetation, basecase (fire/no fire)

```{r}

l1_model_statsums %>% 
  select(short_scenario, scenario, scenario_group, description, scenario_raw, description, climate, land_use, hydrology, slope, percent, severity, doc_perc_mean, doc_perc_median,  doc_perc_sd, nitrate_perc_mean, nitrate_perc_median, nitrate_perc_sd, sed_perc_mean, sed_perc_median, sed_perc_sd) %>%
  filter(scenario_group %in% c("Climate", "Base Case", "Land Use")) %>%
  mutate(scenario = case_when(scenario == 'base_burned' ~ 'burned',
                              scenario == 'WET' ~ 'increased precip',
                              scenario == 'DRY' ~ 'decreased precip',
                              TRUE ~ scenario)) %>%
  ggplot(aes(doc_perc_mean, scenario, color = scenario),
              position = position_dodge(width = -0.5)) +
  geom_pointrange(aes(xmin = doc_perc_mean - doc_perc_sd, xmax = doc_perc_mean + doc_perc_sd, 
                      color = scenario),
                      position = position_dodge(width = -0.5), size = 0.5) +
  geom_vline(xintercept = 0, linewidth = 0.5, color = "red") +
  facet_wrap(~scenario_group, ncol= 1, scales = "free_y") +
  cowplot::theme_cowplot() +
  theme(legend.position = "bottom")

```

